Models of VOR learning in MHC knockout mice

### 2013-01-25 Notes/VORmodel.tex
added pooled resource model

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\tpre}{t_{\text{pre}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Models of VOR learning in MHC knockout mice}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

\begin{abstract}
  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild type.
\end{abstract}

@@

We make the following assumptions:
\begin{itemize}
  \item There are $N$ identical synapses with $M$ internal functional states.
  \item The states of different synapses are independent of each other.
  \item The synapses that are eligible for plasticity are chosen randomly.
  \item The potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
  \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take two values that we can call $\pm1$.
\end{itemize}

@@

\begin{table}
 \begin{center}
  \begin{tabular}{|l|c|c|c|c|c|c|}
    \hline
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $rt_\text{train}$ & $rt_\text{pre}$ \\
    \hline
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50 \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50 \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20 \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5 & 5 \\
    \hline
  \end{tabular}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

@@

\subsubsection{Cascade}\label{sec:cascade}

@@

\subsubsection{Multistate}\label{sec:multistate}

@@

\subsubsection{Two-state}\label{sec:binary}

@@

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
\input{sl_graphics_preamble.tex}
\graphicspath{{"Figures/"}}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\pr}{\mathbf{p}}
\newcommand{\eq}{\pr^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\frg}{\W^\mathrm{F}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
\newcommand{\pot}{^{\text{pot}}}
\newcommand{\dep}{^{\text{dep}}}
\newcommand{\potdep}{^{\text{pot/dep}}}
\newcommand{\norm}{_0}
\newcommand{\inc}{_{\text{inc}}}
\newcommand{\dec}{_{\text{dec}}}
\newcommand{\incdec}{_{\text{inc/dec}}}
\newcommand{\wt}{_{\text{WT}}}
\newcommand{\ko}{_{\text{D$^\mathrm{b}$K$^\mathrm{b}$-/-}}}
\newcommand{\KO}{D$^\mathrm{b}$K$^\mathrm{b}$-/-}
\newcommand{\tpre}{t_{\text{pre}}}
\newcommand{\ttrain}{t_{\text{train}}}
\newcommand{\lmax}{_{\text{max}}}
\newcommand{\lmin}{_{\text{min}}}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Models of VOR learning in MHC knockout mice}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

\begin{abstract}
  We see if we can model VOR gain increase and decrease learning in mice with a knockout in MHC as well as wild-type.
\end{abstract}

@@

We make the following assumptions:
\begin{itemize}
  \item There are $N$ identical synapses with $M$ internal functional states.
  \item The states of different synapses are independent of each other.
  \item The synapses that are eligible for plasticity are chosen randomly.
  \item The potentiating/depressing plasticity event timings are distributed as Poisson processes with rates $rf\potdep$, where $f\pot+f\dep=1$.
  \item Potentiation and depression are described by Markov processes with transition probabilities $\M\potdep$.
  \item The synaptic weights of the internal states are given by the column vector $\w$. This can only take two values that we can call $\pm1$, except for the pooled resource model that we will discuss in \sref{sec:pooledmodel} and \sref{sec:pooled}.
\end{itemize}

@@

\begin{table}
 \begin{center}
  \begin{tabular}{|l|c|c|c|c|c|c|}
    \hline
    % after \\: \hline or \cline{col1-col2} \cline{col3-col4} ...
    Model & \# states & pot,WT dep & \KO\ dep & $\Delta f$ & $r\ttrain$ & $r\tpre$ \\
    \hline
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 50 \\
    Cascade    & 10 & $x=0.25$ & $x=0.33$ & -0.3 & 50 & 150 \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.1 & 50 & 50 \\
    Multistate & 10 & $q=0.6$  & $q=0.8$  & -0.4 & 20 & 20 \\
    Two-state  & 2  & $q=0.4$  & $q=0.8$  & -0.1 & 5 & 5 \\
    \hline
  \end{tabular}
 \end{center}
  \caption{Parameters used in simulations.} \label{tab:params}
\end{table}

\subsubsection{Pooled resource model}\label{sec:pooledmodel}

Suppose that there is some resource required for potentiation/depression that is shared between $P$ synapses and is depleted as more synapses are potentiated/depressed.
We can avoid going beyond the independent synapse model by modelling this pool of synapse as a single pseudo-synapse.

We will model the individual synapses with the two-state model. 
Let $i$ be the number of them that are potentiated.
We will model the effect of resource depletion linearly with the potentiation/depression probabilities for the individual synapses:
%
\begin{equation}\label{eq:depletion}
  \begin{aligned}
    q\pot &= \frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}, \qquad
    q\dep &= \frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}.
  \end{aligned}
\end{equation}
%
At each plasticity event for the pseudo-synapse, one of the individual synapses will be chosen randomly for update.
This effectively reduces the transition probabilities by $1/P$.

This pseudo-synapse would seem to have $2^P$ internal states.
However, we need only keep track of the number of potentiated synapses, not their identity, leaving $M=P+1$ states.
The transition network topology will then have a multistate toplogy (see \autoref{fig:models}\ref{fig:multistate}) but the transition probabilities will no longer be uniform and the weight of the pseudo-synapse is the mean of its constituent synapses:
%
\begin{equation}\label{eq:pooledweight}
  \w_i = \frac{2i}{P}-1.
\end{equation}
%

The Markov process is lumpable \wrt this partition of states (see \cite[\S6.3]{kemeny1960finite} for the discrete time version and \cite{burke1958markovian,Ball1993Lumpability} for continuous time).
The transition probabilities between lumps $i$ and $j$ is computed by choosing one state from lump $i$ and summing the transition probabilities to all states in lump $j$.
The result must be the same for all states in lump $i$.

For any state in lump $i$, there are $P-i$ synapses that can be potentiated to go to lump $i+1$.
Each of these transition probabilities is $q\pot/P$.
Similarly, there are $i$ synapses that can be depressed to go to lump $i-1$.
Each of these transition probabilities is $q\dep/P$.
Thus:
%
\begin{equation}\label{eq:pooledpotdep}
  \begin{aligned}
    \M\pot_{ii+1} &=  \brk{\frac{(P-i-1)q\lmax + i\,q\lmin}{P-1}} \frac{P-i}{P}, \\
    \M\dep_{ii-1} &=  \brk{\frac{(i-1)q\lmax + (P-i)q\lmin}{P-1}} \frac{i}{P},
  \end{aligned}
\end{equation}
%
with all other off-diagonal elements equal to zero.
The diagonal elements are chosen so that the rows sum to one.

@@

\subsubsection{Cascade model}\label{sec:cascade}

@@

\subsubsection{Multistate model}\label{sec:multistate}

@@

\subsubsection{Two-state model}\label{sec:binary}

@@

\subsubsection{Pooled resource model}\label{sec:pooled}
